---
title: "Coursera RR Assignment 1"
author: "Tomasz Kun"
date: "09 oct 2018"
output:
  html_document:
      keep_md: yes
  pdf_document: default
---

### Prepare Local Environment

```{r Prepare Local Environment, results = "hide", message = FALSE, warning = FALSE, cache = TRUE}

Sys.setlocale("LC_MESSAGES", 'en_GB.UTF-8')
Sys.setenv(LANG = "en_US.UTF-8")

#setwd("C:/Users/tomas/OneDrive/Dokumenty/R/Coursera/ReproducibleResearch/Assignment1")

if (!dir.exists("LandingZone")){
  dir.create("LandingZone")
  print("Folder LandingZone created!")
} else {
  print("Folder LandingZone already exists!")
}

if (!dir.exists("Data")){
  dir.create("Data")
  print("Folder Data created!")
} else {
  print("Folder Data already exists!")
}


download.file( url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",  
               destfile = "LandingZone/ActivityMonitoringData.zip", 
               method = "libcurl" )

unzip( zipfile = "LandingZone/ActivityMonitoringData.zip", exdir = "Data" )

```


### 1. Code for reading in the dataset and/or processing the data

```{r Load the data, results='hide', cache = TRUE}

d0 <- read.csv("Data/activity.csv", header = TRUE, sep = ",", na.strings= "NA", 
         colClasses = c("numeric", "character", "numeric"))

d0$date     <- as.POSIXct( strptime(d0$date, "%Y-%m-%d" ))

```

 
 
### 2. Histogram of the total number of steps taken each day

#### Total number of steps per day

```{r, results = "hide", message = FALSE, warning = FALSE}

#install.packages("dplyr")
library("dplyr")

```

```{r}

a1 <- d0 %>% 
  group_by(date) %>%
  summarise(steps = sum(steps))
 
with(a1, barplot(steps, names.arg = format(a1$date, format="%d.%m"), main = "No. Steps by Date (2018)"))
 

```

#### Histogram and Boxplot
```{r}
par(mfrow = c(1,2))
with(a1, hist(x = steps, xlab = ""))
with(a1, boxplot(x = steps, main = "Boxplot of steps"))
```


### 3. Mean and median number of steps taken each day
```{r}
mean(a1$steps, na.rm = TRUE)
median(a1$steps, na.rm = TRUE)
```


### 4. Time series plot of the average number of steps taken
```{r}

a2 <- d0 %>% 
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm = TRUE))

mean_a2_steps <- mean(a2$steps)

with(a2, plot(interval, steps, type = "l", main = "Average Number of Steps Taken"))
abline(h = mean_a2_steps, col = "gray60")
text(x = 0, y = mean_a2_steps+5, format(round(mean_a2_steps, 2), nsmall = 2), col = 1, adj = c(-.1, -.1))

```

 
### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}

subset(a2, steps == max(steps))
```

 
### 6. Code to describe and show a strategy for imputing missing data
#### Investigate Null values
```{r}

# 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

summary(d0)
x1 <- table(is.na(d0))
x2 <- matrix(x1, nrow = 1, ncol = 2, dimnames = list(c("no. occurencies"),c("Non NULL","NULL")))

# number of NULL occurencies
x2

# Percentage of NULL occurencies:
x3 <- x2[,2]/x2[,1]

```
There are `r x2[,2]` NULL occurencies representing `r format(x3*100, digits = 2)`% of the total number of rows

#### Calculate mean of a day and a mean for the whole dataset if the day doesnt contain any data.

```{r}

# Calcualte mean of steps for each day
imputed_date_d0 <- d0 %>% 
  group_by(date) %>%
  summarise(steps = mean(steps, na.rm = TRUE))

# For dates with missing data impute them with a mean of steps from all dates
imputed_date_d0 <- imputed_date_d0
imputed_date_d0$steps[is.na(imputed_date_d0$steps) == TRUE] <- mean(d0$steps, na.rm = TRUE)

```

#### Impute missing values in the original dataset 

```{r}

# Use those values to impute missing values in the base dataset
imputed_d0 <- d0

imputed_d0 <- left_join(d0, imputed_date_d0, by = "date")

imputed_d1 <- rename(imputed_d0, 
       steps_src = steps.x,
       steps_imp = steps.y )

imputed_d1$steps <- coalesce(imputed_d1$steps_src,imputed_d1$steps_imp )

imputed_data <- select(imputed_d1, 
                     date,
                     interval,
                     steps_src,
                     steps_imp,
                     steps)
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}

e1 <- imputed_data %>% 
  group_by(date) %>%
  summarise(steps = sum(steps))
 
with(e1, hist(steps, main = "No. Steps by Date in 2018 (Imputed missing values"))

```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

imputed_data$weekday <- weekdays(imputed_data$date)

weekdays_dataset <- subset(imputed_data, weekday != "Saturday" | weekday != "Sunday" )
weekend_dataset <- subset(imputed_data, weekday == "Saturday" | weekday == "Sunday" )

weekdays_d1 <- weekdays_dataset %>% 
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm = TRUE))

weekend_d1 <- weekend_dataset %>% 
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm = TRUE))

mean_weekdays_d1 <- mean(weekdays_d1$steps)
mean_weekend_d1 <- mean(weekend_d1$steps)


par(mfrow = c(1,2))

with(weekdays_d1, plot(interval, steps, type = "l", main = "Weekdays"))
abline(h = mean_weekdays_d1, col = "gray60")
text(x = 0, y = mean_weekdays_d1+5, format(round(mean_weekdays_d1, 2), nsmall = 2), col = 1, adj = c(-.1, -.1))

with(weekend_d1, plot(interval, steps, type = "l", main = "Weekend"))
abline(h = mean_weekend_d1, col = "gray60")
text(x = 0, y = mean_weekend_d1+5, format(round(mean_weekend_d1, 2), nsmall = 2), col = 1, adj = c(-.1, -.1))


```